|

### 2023-11-06 talks/ISF_talk/talk.qmd

::: columns

::: {.column width="50%"}

About me:

- PhD in pure math
- transitioned to public health data science 
- started as RA with Nick and Evan at [reichlab.io](reichlab.io) at UMass in 2021
- maintain COVID-19 Forecast Hub ensemble
- current research goals center on ensemble methodology
  - focus on scoring rules, decision theory, forecast-policy alignment

:::

::: {.column width="50%"}

![](Reichlab.png)

![](C19FH.png)

:::

:::

@@

![](hub_timeline.png){width=80%}

::: {.attribution}
Figure credits: Alex Vespignani and Nicole Samay
:::

New directions: "Hubverse" infrastructure project [https://hubdocs.readthedocs.io/](), genomic targets

@@

::: box

Create scenario where the forecaster believes maximizing her own expected utility aligns 
with leading the DM toward actions that maximize social utility.

:::

@@

::: box

The **Bayes act** relative to a forecast $Y \sim F$ is the action $x^F$ 
by which a DM will minimize **expected loss** $E_Y[l]$ 
assuming that in reality $Y \sim F$.

:::

 - loss $l$ is a function of the outcome $y$ **and** the action $x$ DM takes in anticipation of $y$

::: box

Key point: Given $l$, we have an *automatically proper* scoring rule 
$S(F,y) := l(x^F, y)$

:::

@@

$$
\begin{align}
l(x,y) &= \sum_{i=1}^{52} \max(y_i - x_i, 0) = \sum (y_i - x_i)_{+}\\
x^F &\in \left\{ \text{possible } x' \,\middle| \, E_{F}[l(x',Y)] \leq E_{F}[l(x,Y)] \text{ for any possible } x \right\}
\end{align}
$$

@@

::: box

Define set of **possible** $x$ as those actions which satisfy $\displaystyle{\sum x_i \leq K}$

:::

@@

\begin{align}
L(x, \lambda ; K, F) &= \sum_{i=1}^N E_{F_i}\left[l(x_i, Y_i)]\right)+\lambda\left(\sum_{i=1}^N x_i-K\right) \\

@@

\nabla_{\lambda} L(x^F, \lambda^{\star}) = 0 
&\implies \sum_{i=1}^{N} F_{i}^{-1}(1-\lambda^{\star}) - K  = 0
\implies 
\boxed{x_{i}^F = F_i^{-1}(1-\lambda^{\star})}
\end{align}

Our allocation scoring rule is then given by the recipe:

@@

So we resort to a binary search for $\lambda^{\star}$ following long history in inventory management of solving **constrained multi-product newsvendor** 
problems^[See e.g., Zhang, Bin, Xiaoyan Xu, and Zhongsheng Hua. 2009. “A Binary Solution Method for the Multi-Product
Newsboy Problem with Budget Constraint.”]

@@

::: columns

::: {.column width="30%"}

After adjusting for an oracle the score of an oracle that allocates *fairly* based on what demands actually occur, scores resemble "value" plots from meteorology

@@

- ranking at peak seems to be consistent with that in large $K$
region.

:::

::: {.column width="70%"}

![](plots/peak_alloscore.png)

:::

:::

::: aside
Ehm, Werner, Tilmann Gneiting, Alexander Jordan, and Fabian Krüger. 2016. “Of Quantiles and Expectiles: Consistent Scoring Functions, Choquet Representations and Forecast Rankings.”
:::

@@

- an allocation score (on joint forecast)
- any allocation based loss
- our quantile interpolation/extrapolation methods (`distfromq`)

@@

[https://github.com/aaronger/alloscore](https://github.com/aaronger/alloscore)

@@

[https://github.com/reichlab/distfromq](https://github.com/reichlab/distfromq)

<!-- ## Loss vs Utility

@@

Change in utility but not in loss from change in - supply and personnel cost structures - public opinion regarding importance of fatality prevention - knowledge of demographic disparities in consequences of outcomes -->

## Data workflow

@@

![](plots/state_hosps_ens_only.jpeg){width=100%}

>>>>

::: columns
::: {.column width="50%"}
About me:

After working as a post-doc in pure math for several years I transitioned to public health data science, joining Nick Reich and Evan Ray's group (reichlab.io) at UMass in 2021 as an RA. I help maintain the COVID-19 Forecast Hub ensemble forecasting system and do research on ensemble methodology, scoring rules, decision theory, and forecast-policy alignment.
:::

::: {.column width="50%"}
![](Reichlab.png)

![](C19FH.png)
:::
:::

@@

![](hub_timeline.png){width="80%"}

```{=html}
<!-- ::: attribution
Figure credits: Alex Vespignani and Nicole Samay
::: -->
```
New directions: "Hubverse" infrastructure project [https://hubdocs.readthedocs.io/](), genomic targets

@@

::: box
Create scenario where the forecaster believes maximizing her own expected utility aligns with leading the DM toward actions that maximize social utility.
:::

@@

::: box
The **Bayes act** relative to a forecast $Y \sim F$ is the action $x^F$ by which a DM will minimize **expected loss** $E_Y[l]$ given that $Y \sim F$.
:::

-   loss $l(x,y)$ is the **consequence** of experiencing $y$ having prepared with $x$

::: box
Key point: Given $l$, we have an *automatically proper* scoring rule $S(F,y) := l(x^F, y)$
:::

@@

```{=tex}
\begin{align*}
l(x,y) &= \sum_{i=1}^{52} \max(y_i - x_i, 0) = \sum (y_i - x_i)_{+}\\
x^F &\in \left\{ \text{possible } x' \,\middle| \, E_{F}[l(x',Y)] \leq E_{F}[l(x,Y)] \text{ for any possible } x \right\}
\end{align*}
```

@@

::: box
Define set of **possible** $x$ as those actions which satisfy $\displaystyle{\sum x_i \leq K}$
:::

@@

```{=tex}
\begin{align}
L(x, \lambda ; K, F) &= \sum_{i=1}^N E_{F_i}\left[l(x_i, Y_i)]\right)+\lambda\left(\sum_{i=1}^N x_i-K\right) \\

@@

\nabla_{\lambda} L(x^F, \lambda^{\star}) = 0 
&\implies \sum_{i=1}^{N} F_{i}^{-1}(1-\lambda^{\star}) - K  = 0
\implies 
\boxed{x_{i}^F = F_i^{-1}(1-\lambda^{\star})}
\end{align}
```
Our allocation scoring rule is then given by the recipe:

@@

So we resort to a binary search for $\lambda^{\star}$ following long history in inventory management of solving **constrained multi-product newsvendor** problems[^1]

[^1]: See e.g., Zhang, Bin, Xiaoyan Xu, and Zhongsheng Hua. 2009. "A Binary Solution Method for the Multi-Product Newsboy Problem with Budget Constraint."

@@

::: columns
::: {.column width="30%"}
After adjusting for an oracle the score of an oracle that allocates *fairly* based on what demands actually occur, scores resemble "value" plots from meteorology

@@

-   ranking at peak seems to be consistent with that in large $K$ region.
:::

::: {.column width="70%"}
![](plots/peak_alloscore.png)
:::
:::

::: aside
Ehm, Werner, Tilmann Gneiting, Alexander Jordan, and Fabian Krüger. 2016. "Of Quantiles and Expectiles: Consistent Scoring Functions, Choquet Representations and Forecast Rankings."
:::

@@

-   an allocation score (on joint forecast)

-   any allocation based loss

-   our quantile interpolation/extrapolation methods (`distfromq`)

@@

<https://github.com/aaronger/alloscore>

@@

<https://github.com/reichlab/distfromq>

```{=html}
<!-- ## Loss vs Utility

@@

Change in utility but not in loss from change in - supply and personnel cost structures - public opinion regarding importance of fatality prevention - knowledge of demographic disparities in consequences of outcomes -->
```
## Data workflow

@@

![](plots/state_hosps_ens_only.jpeg){width="100%"}
